Scheduling and Amortization for MPC

### 2021-12-20 paper_SIMD.tex
Edits to Extension of Basic Vectorization.

Consider a loop $j$ enclosed in some fixed $\underline{i}$. Only if an update (definition) $\texttt{A}_m[f(i,j,k)] = ... $ at some iteration $\underline{j}$ 
references the \emph{same} array element as a use $ ... = \texttt{A}_n[f'(i,j,k)]$ at some later iteration $\underline{j}'$, 
we may have a loop-carried dependence for $\texttt{A}$ due to this def-use pair. (In contrast, Cytron's algorithm inserts a loop-carried dependency every time there is an array update.)
The algorithm above examines all def-use pairs in loop $j$, including defs and uses in nested loops, searching for values $\underline{i}, \underline{j}, \underline{j}', \underline{k}, \underline{k}'$ that satisfy 
$f(\underline{i},\underline{j},\underline{k}) = f'(\underline{i},\underline{j}',\underline{k}')$. If such values exist for some def-use pair, then there is a potential 
loop-carried dependence on $\texttt{A}$; otherwise there is not and we can remove the spurious backward edge thus ``freeing up'' statements for vectorization. \ana{Double check this!!!}

@@

\subsubsection{Array rebase}

\ana{TODO, add here.}

@@

Next, the algorithm refines array MUX statement. 
MPC-source after Cytron's SSA may result in statements $\texttt{A}_j = \emph{MUX}(..., \texttt{A}_k, \texttt{A}_l)$, which imply that any index 
of A can be written at this point and therefore there is a loop-carried dependency. In some cases the MUX can be refined to just a single 
index or a pair of indices, e.g., $\texttt{A}_j[i] = \emph{MUX}(c, \texttt{A}_k[i], \texttt{A}_l[i])$. \ana{Clarify why this helps. Still work in progress.}

\begin{algorithmic}
\FOR {each $\mathit{stmt}$: $\texttt{A}_j = \emph{MUX}(c, \texttt{A}_k, \texttt{A}_l)$ in the MPC-source seq.}
\STATE $i_1 = \mathit{find\_update}(A_k)$ \COMMENT { Is null when $A_k = \phi(...)$ }
\STATE $i_2 = \mathit{find\_update}(A_l)$ \COMMENT { Is null when $A_l = \phi(...)$ }
\IF {$i_1 \cong i_2$} \STATE \COMMENT{ Also $i_1 \cong i_2$ when one is null. WLG assume $i_2$ is null } 
\STATE replace $\mathit{stmt}$ with \\
\STATE $\texttt{A}_j = \texttt{A}_k$; $\texttt{A}_j[i_1] = \emph{MUX}(c, \texttt{A}_k[i_1], \texttt{A}_l[i_1])$
%\ELSIF {$i_1 \ncong i_2$}  
%\STATE replace $\mathit{stmt}$ with \\
%\STATE $\texttt{A}_j = \texttt{A}_k$; $\texttt{A}_j[i_1] = \emph{MUX}(c, \texttt{A}_k[i_1], \texttt{A}_l[i_1])$
\ELSE \STATE \COMMENT{ Also, not $\cong$ when one is $\texttt{A}_j = \emph{MUX}(..., \texttt{A}_k, \texttt{A}_l)$}
\STATE $\mathit{stmt}$ stays as is
\ENDIF
\ENDFOR
\end{algorithmic}

>>>>

Consider a loop $j$ enclosed in some fixed $\underline{i}$. Only if an update (definition) $\texttt{A}_m[f(i,j,k)] = ... $ at some iteration $\underline{j}$ 
references the \emph{same} array element as a use $ ... = \texttt{A}_n[f'(i,j,k)]$ at some later iteration $\underline{j}'$, 
we may have a loop-carried dependence for $\texttt{A}$ due to this def-use pair. (In contrast, Cytron's algorithm inserts a loop-carried dependency every time there is an array update.)
The algorithm above examines all def-use pairs in loop $j$, including defs and uses in nested loops, searching for values $\underline{i}, \underline{j}, \underline{j}', \underline{k}, \underline{k}'$ that satisfy 
$f(\underline{i},\underline{j},\underline{k}) = f'(\underline{i},\underline{j}',\underline{k}')$. If such values exist for some def-use pair, then there is a potential 
loop-carried dependence on $\texttt{A}$; otherwise there is not and we can remove the spurious backward edge thus ``freeing up'' statements for vectorization. %\ana{Double check this!!!}

@@

\ana{TODO: We may remove phi-nodes that end up without incoming back-edges. This can free some cycles that we'd waste on bookkeeping arrays.}

@@

\ana{TODO: I think we should implement this.}

Next, the algorithm refines array MUX statements. 
MPC-source after Cytron's SSA may result in statements $\texttt{A}_j = \emph{MUX}(..., \texttt{A}_k, \texttt{A}_l)$, which imply that any index 
of A can be written at this point and therefore there is a loop-carried dependency. In some cases the MUX can be refined to just a single 
index or a pair of indices, e.g., $\texttt{A}_j[i] = \emph{MUX}(c, \texttt{A}_k[i], \texttt{A}_l[i])$.

%\ana{Clarify why this helps. Still work in progress.}
This is to reduce the dimensionality of simd-ified computation. Technically, $\texttt{A}_j = \emph{MUX}(..., \texttt{A}_k, \texttt{A}_l)$ is 
a simdified operation that can be carried out in parallel "in one round". However, if \texttt{A} is a multi-dimensional array, this would lead 
to substantial increase in the size of the arrays (vectors) we send to SIMD operations. Refining to an update to a specific index would reduce 
the size of those vectors. Note that this is a heuristic that handle a common case, but not all cases of array updates.

\begin{algorithmic}
\FOR {each $\mathit{stmt}$: $\texttt{A}_j = \emph{MUX}(c, \texttt{A}_k, \texttt{A}_l)$ in the MPC-source seq.}
\STATE $i_1 = \mathit{find\_update}(A_k)$ \COMMENT { Is null when $A_k = \phi(...)$ }
\STATE $i_2 = \mathit{find\_update}(A_l)$ \COMMENT { Is null when $A_l = \phi(...)$ }
\IF {$i_1 == i_2$ or $i_1$ is null or $i_2$ is null} \STATE \COMMENT{ With our restrictions on writes we must have $i_1 = i_2$. } 
\STATE replace $\mathit{stmt}$ with \\
\STATE $\texttt{A}_j = \texttt{A}_k$; $\texttt{A}_j[i_1] = \emph{MUX}(c, \texttt{A}_k[i_1], \texttt{A}_l[i_1])$
%\ELSIF {$i_1 \ncong i_2$}  
%\STATE replace $\mathit{stmt}$ with \\
%\STATE $\texttt{A}_j = \texttt{A}_k$; $\texttt{A}_j[i_1] = \emph{MUX}(c, \texttt{A}_k[i_1], \texttt{A}_l[i_1])$
\ELSE \STATE
\STATE $\mathit{stmt}$ stays as is
\ENDIF
\ENDFOR
\end{algorithmic}

\subsubsection{Extending Basic Vectorization with Array Writes}

\ana{TODO, add the weekend notes here. The big things are 1) Restricting array writes to "canonical writes" and 2) change of index, e.g., if accessing A[i] in loop j, have to change to accessing A[i,j].}
